---
title: "191010_DCC_paper_analysis_followalong"
author: "Evan Witt"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Welcome to the source code for Witt and Zhao 2019/2020.  In order to follow along you will need to obtain the data files from the included box link.  This includes the integrated dataset for the two *D. melanogaster* strains, lists of X, Y and testis specific genes, counts from X, Y and autosomes for every gene in every cell, and coordinates for msl chromatin entry sites.

First, let's set everything up.  If you downloaded the data files somewhere else, set your working directory appropriately.

```{r fig.width=12, message=FALSE, warning=FALSE}
setwd("~/Downloads/WittZhao2020data")
library(Seurat)
library(viridis)
library(MASS)
library(plyr)
library(dplyr)
library(sctransform)
library(hues)
library(ggpubr)
library(ggsignif)
library(tidyverse)
library(rstatix)
library(cowplot)

#This is for times we want to represent p values as asterixis
symnum.args <- list(
  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), 
  symbols = c("****", "***", "**", "*", "")
)

#read in the tesis RDS, and gene lists
testis.integrated<-readRDS("~/Downloads/WittZhao2020data/190730_R517_wild_integrated_SCtransform.RDS")
testis.integrated@active.ident<-factor(testis.integrated@active.ident, levels=c("Hub cells", "Cyst cells", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids"))

testisspecificgenes<-read.csv( "~/Downloads/WittZhao2020data/190516_testisspecificgenes_flyatlas.csv")
names(testisspecificgenes)<-c("row","ID")
genestsv<-read.table("~/Downloads/WittZhao2020data/genes.tsv")
names(genestsv)<-c("ID", "Gene")
X <-(scan("~/Downloads/WittZhao2020data/X.txt", what="", sep="\n"))
X<-X[X %in% rownames(GetAssayData(object = testis.integrated))]

Y <-(scan("~/Downloads/WittZhao2020data/ChrY.txt", what="", sep="\n"))
Y<-Y[Y %in% rownames(GetAssayData(object = testis.integrated))]
autosomal <-rownames(GetAssayData(object = testis.integrated))
autosomal <-autosomal[!autosomal %in% X]
autosomal <-autosomal[!autosomal %in% Y]

testisspecificgenes<-join(testisspecificgenes,genestsv, by="ID", type="inner" )
testisspecificgenes<-droplevels(testisspecificgenes$Gene)
testisspecificgenes<-testisspecificgenes[testisspecificgenes %in% rownames(GetAssayData(object = testis.integrated))]

```


Make a dotplot of all the marker genes in every cell type, then get the barcodes of each cell of every type.  Then get the raw counts from every gene, divided by cell type.
```{r, fig.width=14}
DotPlot(testis.integrated, assay = "RNA", features = rev(c("MtnA","zfh1","Fas3", "aub", "bam", "His2Av", "fzo", "twe", "Dpy-30L2", "p-cup")), cols=c("lightblue", "red"), col.min = 0)

#barcodes
Hub <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Hub cells"))
Cyst <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Cyst cells"))
Epi <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Epithelial cells"))
GSC.Early.spermatogonia <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="GSC, Early spermatogonia"))
Late.spermatogonia <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Late spermatogonia"))
Early.spermatocytes <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Early spermatocytes"))
Late.spermatocytes <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Late spermatocytes"))
Early.spermatids <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Early spermatids"))
Late.spermatids <-as.matrix(subset(Idents(testis.integrated), Idents(testis.integrated)=="Late spermatids"))

#raw counts
Hub.raw<-as.matrix(testis.integrated@assays$RNA@counts[,rownames(Hub)])
Cyst.raw<-as.matrix(testis.integrated@assays$RNA@counts[,rownames(Cyst)])
Epithelial.raw<-as.matrix(testis.integrated@assays$RNA@counts[,rownames(Epi)])
GSC.Early.spermatogonia.raw<-as.matrix(GetAssayData(object=testis.integrated, slot="counts")[,rownames(GSC.Early.spermatogonia)])
Late.spermatogonia.raw<-as.matrix(GetAssayData(object=testis.integrated, slot="counts")[,rownames(Late.spermatogonia )])
Early.spermatocytes.raw<-as.matrix(GetAssayData(object=testis.integrated, slot="counts")[,rownames(Early.spermatocytes)])
Late.spermatocytes.raw<-as.matrix(GetAssayData(object=testis.integrated, slot="counts")[,rownames(Late.spermatocytes)])
Early.spermatids.raw<-as.matrix(GetAssayData(object=testis.integrated, slot="counts")[,rownames(Early.spermatids)])
Late.spermatids.raw<-as.matrix(GetAssayData(object=testis.integrated, slot="counts")[,rownames(Late.spermatids)])


```

Ok let's make figure 1.
```{r, fig.width=10}
#get proportions of expressed genes in every cell type
celltypeprops<-data.frame()
for (i in list("Hub.raw","Cyst.raw", "Epithelial.raw", "GSC.Early.spermatogonia.raw", "Late.spermatogonia.raw", "Early.spermatocytes.raw", "Late.spermatocytes.raw", "Early.spermatids.raw", "Late.spermatids.raw")){
  tmp<-data.frame(colSums(get(i)[X,] > 0)/nrow(get(i)[X,] ), i, "Chr X genes")
  names(tmp)<-c("Proportion", "Celltype", "Genegroup")
  tmp$barcode<-rownames(tmp)
  celltypeprops<-rbind(celltypeprops, tmp)
  tmp<-data.frame(colSums(get(i)[autosomal,] > 0)/nrow(get(i)[autosomal,] ), i, "Autosomal genes")
  names(tmp)<-c("Proportion", "Celltype", "Genegroup")
  tmp$barcode<-rownames(tmp)
  celltypeprops<-rbind(celltypeprops, tmp)
  tmp<-data.frame(colSums(get(i)[Y,] > 0)/nrow(get(i)[Y,] ), i, "Chr Y genes")
  names(tmp)<-c("Proportion", "Celltype", "Genegroup")
  tmp$barcode<-rownames(tmp)
  celltypeprops<-rbind(celltypeprops, tmp)

}

#replace names
celltypeprops$`Celltype`<-gsub(celltypeprops$`Celltype`, pattern=".raw", replacement="")
celltypeprops$`Celltype`<-gsub(celltypeprops$`Celltype`, pattern=".", replacement=" ", fixed=TRUE)
celltypeprops$`Celltype`<-gsub(celltypeprops$`Celltype`, pattern="GSC", replacement="GSC,", fixed=TRUE)
celltypeprops$`Celltype`<-factor(x=celltypeprops$`Celltype`, levels<-c("Hub","Cyst", "Epithelial", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids"))

#make a color palette for our 9 cell types
pal<-as.vector(iwanthue(n = 9))


#make boxplots  plot of RNA for X, autosomes
Xcounts<-colSums(as.matrix(testis.integrated@assays$RNA@counts)[X,])
Autocounts<-colSums(as.matrix(testis.integrated@assays$RNA@counts)[autosomal,])
testis.integrated<-AddMetaData(testis.integrated,metadata = Xcounts, col.name = "Xcounts")
testis.integrated<-AddMetaData(testis.integrated,metadata = Autocounts, col.name = "Autocounts")

data1<-data.frame(Xcounts=testis.integrated@meta.data$Xcounts, Autocounts=testis.integrated@meta.data$Autocounts, Celltype=testis.integrated@active.ident )

data1$Celltype<-factor(data1$Celltype, levels=rev(c("Hub cells", "Cyst cells", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids")))

#get median RNA counts for every cell type, X and autosome
celltypecounts<-data.frame()
for (i in unique(data1$Celltype)){
  tmp<-data.frame(Celltype=i, Xcounts=median(subset(data1, Celltype==i)$Xcounts), Autocounts=median(subset(data1, Celltype==i)$Autocounts))
  celltypecounts<-rbind(tmp, celltypecounts)
}


#Figure 1B  
p1<-ggplot(data1, aes(x=data1$Celltype, y=Xcounts))+geom_boxplot(outlier.shape = NA, aes(x = Celltype, group=Celltype, fill=Celltype, color=Celltype))+theme(legend.position="none")+scale_color_manual(aesthetics = c("color","fill"),values = rev(pal))+
stat_summary(geom = "crossbar", width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+ scale_y_continuous(trans = 'log10',  breaks = scales::trans_breaks("log10", function(x) 10^x),labels =scales::trans_format("log10", scales::math_format(10^.x)))+xlab("Cell type")+ ylab("ChrX RNA per cell")+coord_flip()+
  theme(axis.text.x=element_text(angle=90, size=10), axis.title=element_text(size=10) ,axis.text.y=element_text(size=10))
  
p2<-ggplot(data1, aes(x=data1$Celltype, y=Autocounts))+geom_boxplot(outlier.shape = NA, aes(x = Celltype, group=Celltype, fill=Celltype, color=Celltype))+theme(legend.position="none")+scale_color_manual(aesthetics = c("color","fill"),values = rev(pal))+
  stat_summary(geom = "crossbar", width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+ scale_y_continuous(trans = 'log10',  breaks = scales::trans_breaks("log10", function(x) 10^x),labels = scales::trans_format("log10", scales::math_format(10^.x)))+xlab("Cell type")+ ylab("Autosomal RNA per cell")+coord_flip()+
  theme(axis.text.x=element_text(angle=90, size=10), axis.title.x=element_text(size=10) ,axis.text.y=element_blank(), axis.title.y=element_blank())

p1<-plot_grid(p1, p2, ncol=2, rel_widths=c(2,1))
#Figure 1A- make t-sNE showing dimensional reduction of cell types
p2 <- DimPlot(testis.integrated, reduction = "tsne", label = FALSE , cols = pal)
p2<-p2+theme(text=element_text(size=10), axis.text=element_text(size=10), axis.title=element_text(size=10), legend.text=element_text(size=10))
p1<-p1+theme(text=element_text(size=10), axis.text=element_text(size=10), axis.title=element_text(size=10), legend.text=element_text(size=10))

p1 <-plot_grid(p2, p1, rel_widths = c(1,1.2), rel_heights = c(0.9,1),labels = c("A", "B"), label_size = 10)

celltypeprops$Genegroup<-factor(celltypeprops$Genegroup, levels=c("Chr X genes", "Autosomal genes","Chr Y genes"))

my_comparisons<-list(c("Chr X genes", "Autosomal genes"))

#figure 1 C
p2<-ggplot(celltypeprops, aes(x=Genegroup, y=Proportion))+  geom_boxplot(outlier.shape=NA, aes_string(colour=celltypeprops$`Genegroup`, fill=celltypeprops$`Genegroup`))+
  stat_summary(geom = "crossbar", width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+
  xlab("Cell type")+ylab('Proportion of genes expressed')+  guides(fill=guide_legend(title="Gene class"))+
  theme( legend.position="top", axis.ticks.x=element_blank(), axis.text.x=element_blank(), text=element_text(size=10))+
  facet_wrap(~`Celltype`, scales="free_x", nrow=1, strip.position = "bottom",labeller = label_wrap_gen(width=15))+
  theme(strip.text=element_text(margin=margin(2,0,2,0, "pt"), size=10), strip.background=element_rect(fill="white"), strip.placement="outside", text=element_text(size=10), axis.text=element_text(size=10), axis.title=element_text(size=10), legend.text=element_text(size=10))+guides(colour=FALSE)

#get raw and adjusted p values for figure 1C
stat.test <- celltypeprops %>%
  group_by(Celltype) %>%
  wilcox_test(Proportion ~ Genegroup, comparisons = my_comparisons, alternative="less") %>%
  adjust_pvalue(method = "holm") %>%
  add_significance("p.adj") %>%
  add_y_position()


p2<-p2+ stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0.01 )
 
plot_grid(p1, p2, ncol=1, labels = c("A", "C"), label_size = 10, rel_heights = c(0.8,1.5))



```
Time for figure 2:
```{r, fig.width=10, fig.height=10, fig.height=6.5}

#Get expression of every gene in every cell
testis.scaled<-as.matrix(GetAssayData(testis.integrated, assay  = "RNA"))
#scale every gene so that 0 is its median expression across all cells
testis.scaled<-t(scale(t(testis.scaled)))
testis.scaled[is.nan(testis.scaled)]<-0

#get scaled expression of cells in each celltype
Hub.scale<-as.matrix(testis.scaled[,rownames(Hub)])
Epi.scale<-as.matrix(testis.scaled[,rownames(Epi)]) 
Cyst.scale<-as.matrix(testis.scaled[,rownames(Cyst)])
GSC.Early.spermatogonia.scale<-as.matrix(testis.scaled[,rownames(GSC.Early.spermatogonia)])
Late.spermatogonia.scale<-as.matrix(testis.scaled[,rownames(Late.spermatogonia )])
Early.spermatocytes.scale<-as.matrix(testis.scaled[,rownames(Early.spermatocytes)])
Late.spermatocytes.scale<-as.matrix(testis.scaled[,rownames(Late.spermatocytes)])
Early.spermatids.scale<-as.matrix(testis.scaled[,rownames(Early.spermatids)])
Late.spermatids.scale<-as.matrix(testis.scaled[,rownames(Late.spermatids)])

#get mean expression of every gene across every celltype
#Somatic.bias <-data.frame(Celltype="Somatic cells", Exp=rowMeans(Somatic.scale), Gene=rownames(Somatic.scale), row.names = NULL)
Hub.bias <-data.frame(Celltype="Hub cells", Exp=rowMeans(Hub.scale), Gene=rownames(Hub.scale), row.names = NULL)
Cyst.bias <-data.frame(Celltype="Cyst cells", Exp=rowMeans(Cyst.scale), Gene=rownames(Cyst.scale), row.names = NULL) 
Epi.bias <-data.frame(Celltype="Epithelial cells", Exp=rowMeans(Epi.scale), Gene=rownames(Epi.scale), row.names = NULL) 
GSC.Early.spermatogonia.bias <-data.frame(Celltype="GSC, Early spermatogonia", Exp=rowMeans(GSC.Early.spermatogonia.scale), Gene=rownames(GSC.Early.spermatogonia.scale), row.names = NULL)
Late.spermatogonia.bias <-data.frame(Celltype="Late spermatogonia", Exp=rowMeans(Late.spermatogonia.scale), Gene=rownames(Late.spermatogonia.scale), row.names = NULL)
Early.spermatocytes.bias <-data.frame(Celltype="Early spermatocytes", Exp=rowMeans(Early.spermatocytes.scale), Gene=rownames(Early.spermatocytes.scale), row.names = NULL)
Late.spermatocytes.bias <-data.frame(Celltype="Late spermatocytes", Exp=rowMeans(Late.spermatocytes.scale), Gene=rownames(Late.spermatocytes.scale), row.names = NULL)
Early.spermatids.bias <-data.frame(Celltype="Early spermatids", Exp=rowMeans(Early.spermatids.scale), Gene=rownames(Early.spermatids.scale), row.names = NULL)
Late.spermatids.bias <-data.frame(Celltype="Late spermatids", Exp=rowMeans(Late.spermatids.scale), Gene=rownames(Late.spermatids.scale), row.names = NULL)

#makae data frame with scaled epxression of every gene in every cell type
plot.data<-rbind(Hub.bias,Cyst.bias,Epi.bias,GSC.Early.spermatogonia.bias,Late.spermatogonia.bias,Early.spermatocytes.bias,Late.spermatocytes.bias,Early.spermatids.bias,Late.spermatids.bias )
data1 <-data.frame(subset(plot.data, Gene %in% X), "Chr X genes")
data3<-data.frame(subset(plot.data, Gene %in% Y), "Chr Y genes")
data2<-data.frame(subset(plot.data, Gene %in% autosomal), "Autosomal")

library(ggpubr)
names(data1)<-c ("Celltype","Exp", "Gene", "Class" )
names(data2)<-c ("Celltype","Exp", "Gene", "Class" )
names(data3)<-c ("Celltype","Exp", "Gene", "Class" )
#data4 has, for every gene, its mean scaled expression across every cell type
data4 <-rbind(data1, data2, data3)
data4$Class<-factor(data4$Class, levels=c("Chr X genes", "Autosomal","Chr Y genes"))


my_comparisons<-list(c("Chr X genes", "Autosomal"), c("Chr Y genes", "Autosomal"))
#test if gene class affects scaled expression in every cell type, adjusts p value with holms correction
stat.test1 <- data4 %>%
  group_by(Celltype) %>%
  wilcox_test(Exp ~ Class, comparisons = my_comparisons, alternative="greater")
stat.test1 <- stat.test1 %>% add_y_position()
stat.test2 <- data4 %>%
  group_by(Celltype) %>%
  wilcox_test(Exp ~ Class, comparisons = my_comparisons, alternative="less")
stat.test2 <- stat.test2 %>% add_y_position()
stat.test1$p.adj<-p.adjust(stat.test1$p)
stat.test2$p.adj<-p.adjust(stat.test2$p)
stat.test1$p.adj.signif<- symnum(stat.test1$p.adj, corr = FALSE,
                                 cutpoints = c(0,.00005,  .0005,.005,.05,  1),
                                 symbols = c("****","***","**","*"," "))
stat.test2$p.adj.signif<- symnum(stat.test2$p.adj, corr = FALSE,
                                 cutpoints = c(0,.00005,  .0005,.005,.05,  1),
                                 symbols = c("****","***","**","*"," "))

stat.test1<-subset(stat.test1, group2 != "Chr Y genes")
stat.test2<-subset(stat.test2, group2 != "Chr Y genes")
#This is to put it in the right place when we plot
stat.test1$y.position<-1.5
stat.test2$y.position<-1.5
#put cute little arrows in to show the relationship between sex chromosome and autosome- up or down
stat.test1$p.adj.signif<-gsub(stat.test1$p.adj.signif, pattern = "^\\*", replacement=paste("\U2191", "*"))
stat.test2$p.adj.signif<-gsub(stat.test2$p.adj.signif, pattern = "^\\*", replacement=paste("\U2193", "*"))
#Don't print NS on plot.  We already have asterixis
stat.test1$p.adj.signif<-gsub(stat.test1$p.adj.signif, pattern = "ns", replacement="")
stat.test2$p.adj.signif<-gsub(stat.test2$p.adj.signif, pattern = "ns", replacement="")
#setting bracket size to 0 puts the p.adj asterixes in between each boxplot.  The rest of the stuff makes the boxpplot, makes it pretty, sets the scale
#First, the X.  Notice that we calculated the statistics first before subsetting the data.
data4<-rbind(data1, data2)
p1<-ggplot(data4, aes(x=Class, y=Exp)) +geom_boxplot(outlier.shape=NA, aes_string(y="Exp", colour="Class", fill="Class"))+facet_wrap(~Celltype, nrow=1,labeller = label_wrap_gen(width=10))+theme_bw()+xlab(NULL) +
  ylab('Scaled expression') +
  stat_summary(geom = "crossbar", width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+
  theme(panel.grid.major = element_blank(),strip.background = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+
   stat_pvalue_manual(stat.test1,remove.bracket = FALSE,bracket.size = 0, label = "p.adj.signif", tip.length = 0.01)+theme(text=element_text(size=10))+
  stat_pvalue_manual(stat.test2,remove.bracket = FALSE,bracket.size = 0,label = "p.adj.signif", tip.length = 0.01)+ylim(-1,2)


p1<-ggplot(data4, aes(x=Class, y=Exp)) +geom_boxplot(outlier.shape=NA, aes_string(y="Exp", colour="Class", fill="Class"))+facet_wrap(~Celltype, nrow=1,labeller = label_wrap_gen(width=10))+theme_bw()+xlab(NULL) +
  ylab('Scaled expression') +
  stat_summary(geom = "crossbar", width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+
  theme(panel.grid.major = element_blank(),strip.background = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  stat_pvalue_manual(stat.test1)
#ok same thing with the Y chromosome
data4 <-rbind(data3,data1, data2)
data4$Class<-factor(data4$Class, levels=c("Chr Y genes", "Autosomal","Chr X genes"))
my_comparisons<-list(c("Chr Y genes", "Autosomal"), c( "Autosomal", "Chr X genes"))

stat.test1 <- data4 %>%
  group_by(Celltype) %>%
  wilcox_test(Exp ~ Class ,ref.group = "Chr Y genes",comparisons = my_comparisons, alternative="greater")
stat.test1 <- stat.test1 %>% add_y_position()
stat.test1$y.position<-0.5
stat.test2 <- data4 %>%
  group_by(Celltype) %>%
  wilcox_test(Exp ~ Class, ref.group="Chr Y genes",comparisons = my_comparisons, alternative="less")
stat.test2 <- stat.test2 %>% add_y_position()
stat.test2$y.position<-0.5
stat.test1$p.adj<-p.adjust(stat.test1$p)
stat.test2$p.adj<-p.adjust(stat.test2$p)
stat.test1$p.adj.signif<- symnum(stat.test1$p.adj, corr = FALSE,
                                 cutpoints = c(0,.00005,  .0005,.005,.05,  1),
                                 symbols = c("****","***","**","*"," "))
stat.test2$p.adj.signif<- symnum(stat.test2$p.adj, corr = FALSE,
                                 cutpoints = c(0,.00005,  .0005,.005,.05,  1),
                                 symbols = c("****","***","**","*"," "))
stat.test1<-subset(stat.test1, group2 != "Chr X genes")
stat.test2<-subset(stat.test2, group2 != "Chr X genes")


stat.test1$p.adj.signif<-gsub(stat.test1$p.adj.signif, pattern = "^\\*", replacement=paste("\U2191", "*"))
stat.test2$p.adj.signif<-gsub(stat.test2$p.adj.signif, pattern = "^\\*", replacement=paste("\U2193", "*"))
stat.test1$p.adj.signif<-gsub(stat.test1$p.adj.signif, pattern = "ns", replacement="")
stat.test2$p.adj.signif<-gsub(stat.test2$p.adj.signif, pattern = "ns", replacement="")

data4<-rbind(data3, data2)
p2<-ggplot(data4, aes(x=Class, y=Exp)) +geom_boxplot(outlier.shape=NA, aes_string(y="Exp", colour="Class", fill="Class"))+facet_wrap(~Celltype, nrow=1,labeller = label_wrap_gen(width=10))+theme_bw()+xlab(NULL) +
  ylab('Scaled expression') +
  stat_summary(geom = "crossbar", width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+
  theme(panel.grid.major = element_blank(),strip.background = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  stat_pvalue_manual(stat.test1,remove.bracket = FALSE,bracket.size = 0,label = "p.adj.signif", tip.length = 0.01)+theme(text=element_text(size=10))+
  stat_pvalue_manual(stat.test2, remove.bracket = FALSE, bracket.size=0,label = "p.adj.signif", tip.length = 0.01)+ylim(-0.6,1)+scale_color_manual(aesthetics = c("color", "fill"), values=c("#C4961A","#00AFBB"))

p3<-plot_grid(p1, p2, nrow=2, labels=c("A", "B"),label_size = 10)
p3<-p3+theme(text=element_text(size=10), axis.text=element_text(size=10))


#encoding problem: arrows look like boxes. If you want to save this one as a pdf, use quartz
levels(testis.integrated)<-rev(c("Hub cells", "Cyst cells", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids"))
p2<-DotPlot(testis.integrated, assay = "RNA", features = rev(c("msl-1", "msl-2", "msl-3", "mle", "mof", "Clamp", "roX1", "roX2")), cols=c("lightblue2",  "red2"), col.min = 0)
p2<-p2+theme(text=element_text(size=10), axis.text.x=element_text(size=10, face = "italic"), axis.text.y=element_text(size=10, face="plain"))+xlab("Genes")+ylab("Cell type")

plot_grid(p3, p2, ncol=1, labels=c("A", "C"), label_size = 10, rel_widths=c(2,1))


#to save this locally, run this:
#quartz(type="pdf", file= "filename.pdf", width=10, height=6.5)
#plot_grid(p3, p2, ncol=1, labels=c("A", "C"), label_size = 10, rel_widths=c(2,1))
#dev.off
```


Oh boy that was fun, let's do figure 3!

This chunk takes hours. Don't run this. To save you some time I put the output of this chunk into a files with the counts for every X and autosomal gene in every cell. Skip to the next part!

```{r eval=F, echo=T}
data1<-data.frame()
for (i in list( "Hub.raw", "Cyst.raw", "Epithelial.raw", "GSC.Early.spermatogonia.raw", "Late.spermatogonia.raw", "Early.spermatocytes.raw", "Late.spermatocytes.raw", "Early.spermatids.raw", "Late.spermatids.raw")){ 
  tmp1<-data.frame(get(i))
  
  for (j in X){
    tmp2<-data.frame(sum(t(tmp1[j,])), j,i, "X")
    names(tmp2)<-c ("Counts", "Genename", "Celltype", "Chr" )
    data1<-rbind(data1, tmp2)
    
  }
  for (j in autosomal){
    tmp2<-data.frame(sum(t(tmp1[j,])), j,i, "Autosomal")
    names(tmp2)<-c ("Counts", "Genename", "Celltype", "Chr" )
    data1<-rbind(data1, tmp2)
    
  }
}

data1$Celltype<-gsub(x=data1$Celltype, pattern="Hub.raw", replacement="Hub cells")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Cyst.raw", replacement="Cyst cells")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Epithelial.raw", replacement="Epithelial cells")
data1$Celltype<-gsub(x=data1$Celltype, pattern="GSC.Early.spermatogonia.raw", replacement="GSC, Early spermatogonia")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Late.spermatogonia.raw", replacement="Late spermatogonia")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Early.spermatocytes.raw", replacement="Early spermatocytes")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Late.spermatocytes.raw", replacement="Late spermatocytes")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Early.spermatids.raw", replacement="Early spermatids")
data1$Celltype<-gsub(x=data1$Celltype, pattern="Late.spermatids.raw", replacement="Late spermatids")
data1$Celltype<-factor(data1$Celltype, levels=rev(c("Hub cells", "Cyst cells", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids")))
write.table(data1, "~/Downloads/WittZhao2020data/190924_X_autosome_counts.txt")



```


You skipped to this part right? Good. Run this chunk, which makes figure 3.

```{r, fig.width=8}
data1<-read.table("~/Downloads/WittZhao2020data/190924_X_autosome_counts.txt")
data1$Celltype<-factor(data1$Celltype, levels=rev(c("Hub cells", "Cyst cells", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids")))

#get raw and adjusted p values
data2<-data1
data2$Counts<-log2(data2$Counts+1)
stat.test1 <- data2 %>%
  group_by(Celltype) %>%
  wilcox_test(Counts ~ Chr ,ref.group = "Autosomal", alternative="greater")%>%
  adjust_pvalue(method = "holm") %>%
  add_significance("p.adj") %>%
  add_y_position()


p1<-ggplot(data1, aes(x=Celltype, y=log2(Counts+1)))+geom_boxplot(aes(fill=Chr, color=Chr), outlier.shape=NA) +  stat_summary(geom = "crossbar",aes(fill=Chr, color=Chr), width=0.6, position= position_dodge(width = 0.75), fatten=0, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })+guides(color=FALSE)+theme_classic()+
  theme(text=element_text(size=10), axis.text=element_text(size=10))+xlab("Cell type")+labs(fill="Chromosome")+ylab("Log2-normalized counts per gene")+stat_pvalue_manual(stat.test1, x="Celltype", label="p.adj.signif")+coord_flip()
p1

```


It's the home stretch! Let's make figure 4 to correlate chromatin entry site distance to expressed counts per gene!
```{r, message=F, warning=F, plot.height=12, plot.width=10, fig.width=10, fig.height=12}

#read in chromatin entry site coordinates
CES<-read.csv("~/Downloads/WittZhao2020data/190927_ChromatinEntrySites_R5.csv", header=TRUE)
CES$R6.start<-as.integer(str_remove_all(CES$R6.start, ","))
CES$R6.end<-as.integer(str_remove_all(CES$R6.start, ","))

#Get start, end, ID, symbol of every dmel gene

flyCDS<-read.table("~/Downloads/WittZhao2020data/dmel-all-r6.15.genes.fb.coords.txt")[,c(1,3,5,6)]
names(flyCDS)<-c("GENEID", "SYMBOL", "start", "end")

#For each gene, get distance from start to closest CES
CESdist<-data.frame()
for (i in flyCDS$SYMBOL[flyCDS$SYMBOL %in% X]){

    dist<-min(abs(subset(flyCDS, SYMBOL==i)$start - CES$R6.start))
    tmp2<-data.frame(Gene=i, distance=dist)
    CESdist<-rbind(CESdist, tmp2)
 
}

#now, for every gene, make a table with counts and distance to closest CES
data2<-read.table("~/Downloads/WittZhao2020data/190924_X_autosome_counts.txt")
data2<-subset(data2, Chr=="X")
data2$Celltype<-factor(data2$Celltype, levels=c("Hub cells", "Cyst cells", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "Early spermatocytes", "Late spermatocytes", "Early spermatids", "Late spermatids"))
CESdist2<-CESdist
CESdist2$Genename<-CESdist2$Gene
data2<-join(data2,CESdist2)


#get raw and adjusted p values for every cell type correlation.
flycor<-data.frame()
for (i in unique(data2$Celltype)){
  res<-cor.test(x = subset(data2, Celltype== i)$distance, y=log2(subset(data2, Celltype== i)$Counts+1), method = "pearson")
  flycor<-rbind(flycor, data.frame(Celltype=i, pvalue=res$p.value, R=res$estimate))
}

flycor$P.adjust<-p.adjust(flycor$pvalue)
flycor$label<-paste("R =",round(flycor$R,digits=3),"\n", "p.adj =", formatC(flycor$P.adjust, format = "e", digits=2), sep= " ")


#this is to label the cell types we care about
Celltypesignif<-c("Hub cells", "", "Epithelial cells", "GSC, Early spermatogonia", "Late spermatogonia", "","","","")

#Figure 4A
p1<-ggplot(data2, aes(x=distance, y=log2(Counts+1)))+geom_point(size=0.2, alpha=0.2)+scale_color_viridis(option="plasma")+geom_smooth(color="steelblue")+
  facet_wrap(~`Celltype`, scales="free")+geom_label(data=flycor, aes(x=Inf, y=Inf, label=label), vjust=1, hjust=1)+
  theme(strip.background = element_blank(),panel.border = element_rect(colour = "black"),strip.text.x=element_text(margin=margin(b=5)))+xlab("Distance (bp) from msl chromatin entry site (CES)")+xlim(0,500000)
#Figure 4B
p2<-ggplot(flycor, aes(x=P.adjust, y=R))+geom_point(aes(color=Celltype))+geom_vline(xintercept=.05, color="red")+ylab("Pearson's R")+xlab("Holm-adjusted p value for Pearson correlation")+ggtitle("Correlation between closest CES (in bp) and RNA per gene")+
  scale_color_manual(aesthetics = c("color","fill"),values = rev(pal))+geom_text(label=Celltypesignif,nudge_x=0.01, hjust=0.25, vjust=-.4)

plot_grid(p1, p2, ncol = 1, rel_heights=c(2,1), labels = c("A", "B"), label_size=10)   


```

Yay you did it! Feel free to tinker with this code however you want- if you want to assign cell types differently, or do an entirely different analysis, go ahead!  There's lots of stories buried in there, just remember to cite Witt and Zhao 2020 when you publish! Thanks for reading!

